---
title: "Assignment_4_EC"
author: "DanielNielsen"
date: "April 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>
# Extra credit assignment for BCB 546X

<br>

## Install the necessary packages

**Note**: If the packages do not install (like rtracklayer for me in Ubuntu), run the following command in bash:

run sudo apt-get install -y libcurl-dev libxml2-dev

<br>

**Download and load biocLite, so it can be used to install other packages**
```{r}
source("https://bioconductor.org/biocLite.R")
library(BiocInstaller)
```

<br>
**Install rtracklayer, GenomicFeatures, BiostringsGenome, and the annotation package for the house mouse**

```{r}
biocLite("rtracklayer")
biocLite("GenomicFeatures")
biocLite("BSgenome.Mmusculus.UCSC.mm10")
biocLite("GRanges")
```
<br>

## Load libraries
**Loads rtracklayer, BiocInstaller, IRanges, GenomicRanges, and house mouse libraries, so their commands and information can be used.**
```{r}
library(rtracklayer)
library(GenomicFeatures)
library(GenomicRanges)
library(Biostrings)
library(BSgenome.Mmusculus.UCSC.mm10)
```
<br>

## Import Mouse Chromosome 38 info using rtracklayer
```{r}
mm_gtf <-import("~/bds-files/chapter-09-working-with-range-data/Mus_musculus.GRCm38.75_chr1.gtf.gz")
colnames(mcols(mm_gtf)) #read the metadata
```

### Find the subset of genes:
```{r}
# list the number of genes with each type of gene
table(mm_gtf$gene_biotype)

#get protein coding genes in chr1_pcg
chr1_pcg <- mm_gtf[mm_gtf$type == "gene" & mm_gtf$gene_biotype == "protein_coding"]

#inspect chr1_pcg, which contains the protein coding genes
summary(width(chr1_pcg))  #mean 60640
length(chr1_pcg)   #1240

#Get the upstream flanking region of chr1_pcg_3kb_up
chr1_pcg_3kb_up <- flank(chr1_pcg, width=3000)
```
<br>

### Retrieving Promoter Sequence
**Explore**
```{r}
#give the genome to a variable
mm_gm <- BSgenome.Mmusculus.UCSC.mm10

#get the organism name
organism(mm_gm)

#get the genome version
provider(mm_gm)

#find out the provider
provider(mm_gm)

#seqinfo() can be used to examine sequence information
seqinfo(mm_gm)

#BSgenome objects can be searched
matchPattern("GGCGCGCC", mm_gm$chr1)

```

<br>

### Remap the chrosomes

**BSgenome and GRanges use different formats**
```{r}
#
all(seqlevels(chr1_pcg_3kb_up)) %in% seqlevels(mm_gm)

#
gr <- GRanges(c("chr1", "chr2"), IRanges(start=c(3, 4), width=10))
seqlevels(gr)

#
seqlevels(gr) <- c("1","2")
seqlevels(gr)

#
seqlevelsStyle(chr1_pcg_3kb_up)

#
seqlevelsStyle(mm_gm)

# 
seqlevelsStyle(chr1_pcg_3kb_up) <- "UCSC"

all(seqlevels(chr1_pcg_3kb_up)) %in% seqlevels(mm_gm)

```

<br>

### Grab sequence regions
```{r}
# 
chr1_3kb_seqs <- getSeq(mm_gm, chr1_pcg_3kb_up)
chr1_3kb_seqs

```

### Write out the promoter regions to a new file
```{r}
writeXStringSet(chr1_3kb_seqs, file="./mm10_chr1_promoters.fasta", format="fasta")
```

